### 2018-02-15 doc/ElectronGas/paper.tex

\documentclass[aps,twocolumn,showpacs,floatfix,nofootinbib,preprintnumbers,superscriptaddress,amsmath,amssymb]{revtex4-1}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{bm}
\usepackage{color}
\usepackage{float}
\usepackage{dcolumn}
\usepackage{multirow} 
\usepackage{subfigure}

\newcommand{\Ho}{\hat{H}^{\rm od}}
\newcommand{\expec}[1]{\langle #1 \rangle}
\DeclareMathOperator\erf{erf}
\DeclareMathOperator\erfc{erfc}

\begin{document}

\title{Many-body approaches to the homogeneous electron gas in two and three dimensions}

\author{Justin G.~Lietz}
\affiliation{National Superconducting
  Cyclotron Laboratory and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}

\author{Sean B.~S.~Miller}
\affiliation{Depatment of Physics, Chalmers Technical University, G\"oteborg, Sweden}

\author{Scott K.~Bogner}
\affiliation{National Superconducting
  Cyclotron Laboratory and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}

\author{Morten Hjorth-Jensen} 
\affiliation{National Superconducting Cyclotron
  Laboratory and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics, University of Oslo, N-0316 Oslo, Norway}

\begin{abstract} 
We present results for the two- and three-dimensional homogeneous
electron gas using first principle methods like coupled cluster
theory and two recently developed methods, the in-medium Similarity Renormalization. In particular, we present
new results for the correlation energy for the two-dimensional
electron gas and analyze our results in terms of finite size effects and basis functions.
\end{abstract}

\pacs{02.70.Ss, 31.15.A-, 31.15.bw, 31.15.V-, 71.10.Ca, 71.15.-m}

\maketitle

\section{Introduction}
A proper understanding of the role of correlations beyond a mean-field
description is of great importance for quantum many-particle
theories. For all possible quantum-mechanical systems, either finite
ones or infinite systems like the homogeneous electron gas in two or
three dimensions or nuclear or neutron star matter, the concept of an
independent particle motion continues to play a fundamental role in
studies of many-particle systems.  Deviations from such an independent
particle, or mean-field, picture have normally been interpreted as a
possible measure of correlations.  The latter are expected to reveal
important features of both the structure and the dynamics of a
many-particle system beyond a mean-field approximation.

The homogeneous electron gas is one of the few examples of a system of many
interacting particles that allows for a solution of the mean-field
Hartree-Fock equations on a closed form.  To first order in the
electron-electron interaction, this applies to ground state properties
like the energy and its pertinent equation of state as well.  The
homogeneus electron gas is a system of electrons that is not
influenced by external forces except by an attraction provided by a
uniform background of ions. These ions give rise to a uniform
background charge.  The ions are stationary and the system as a whole
is neutral.
Irrespective of this simplicity, this system, in both two and
three-dimensions, has eluded a proper description of correlations in
terms of various first principle methods, except perhaps for quantum
Monte Carlo methods. In particular, the diffusion Monte Carlo
calculations of Ceperley {\em et al.}
\cite{ceperley1980,tanatar1989}, are presently still considered as the
best possible benchmarks for the two- and three-dimensional electron
gas.

The electron gas, in 
two or three dimensions is thus interesting as a test-bed for 
electron-electron correlations. The three-dimensional 
electron gas is particularly important as a cornerstone 
of the local-density approximation in density-functional 
theory. In the physical world, systems 
similar to the three-dimensional electron gas can be 
found in, for example, alkali metals and doped 
semiconductors. Two-dimensional electron fluids are 
observed on metal and liquid-helium surfaces, as well as 
at metal-oxide-semiconductor interfaces. However, the Coulomb 
interaction has an infinite range, and therefore 
long-range correlations play an essential role in the
electron gas. At low densities, the electrons become 
localized and form a lattice. This so-called Wigner 
crystallization \cite{wigner1934} is a direct consequence 
of the long-ranged repulsive interaction. At higher
densities, the electron gas is better described as a
liquid \cite{ceperley1978,ceperley1980,tanatar1989,bishop1982}. 
When using, for example, Monte Carlo methods 
\cite{foulkes2001}, the electron gas must be approximated 
by a finite system. The long-range Coulomb interaction 
in the electron gas causes additional finite-size effects 
\cite{fraser1996,chiesa2006,drummond2008} that are not
present in other infinite systems like nuclear matter or neutron star matter in 
\cite{hagen2014}. 
This poses additional challenges to many-body methods when applied 
to the electron gas.

The aim of this work is to present results for the infinite electron
gas using several many-body methods.  We focus in particular on
coupled-cluster theory (add ref), the recently develop in-medium
similarity renormalization group methods, full-configuration
interaction quantmum Monte Carlo \cite{booth2009,shepherd2012a,booth2013}
We present results for all these methods and study in particular both finite size effects and effects due to
a truncation in the single-particle basis (add more refs).

More

\section{Formalism}
\label{sec:formalism}

\subsection{Previous studies of the electron gas}

The electron gas has been studied using a large number 
of different approaches, and we will here only mention
some of the most important works that are relevant here. 
We start with the three-dimensional electron 
gas, which has got most attention in the literature.
The three-dimensional electron gas will mainly serve as a benchmark for our calculations,

It is a well-kown fact that the correlation energy
of the three-dimensional electron gas diverges 
at second order in perturbation theory. The particle-hole ring diagrams 
dominate at the limit of high density. Even though
all these diagrams diverge when calculated separately,
the energy converges when summing all direct 
particle-hole ring diagrams to infinite order.
Gell-Mann and Brueckner obtained for this so-called
random-phase approximation (RPA) the first terms of
the exact energy in the high-density limit 
\cite{gellmann1957}.

At metallic and lower densities, short-ranged correlations
become significant. Singwi \emph{et al.} \cite{singwi1968}
and Lowy and Brown \cite{lowy1975} came up with early attempts 
to combine RPA with contributions for short-ranged 
correlations. The calculations of Singwi \emph{et al.}
were based on dielectric function theory, whereas Lowy and
Brown interpolated between short- and long-range models
using a diagrammatic technique.

In 1978, Ceperley used \cite{ceperley1978} 
the variational Monte Carlo (VMC) method to 
study the two- and three-dimensional electron gas. The
obtained VMC ground-state energies were shown to be close
to other results at that time. Higher accuracy can be obtained with
the diffusion Monte Carlo (DMC) method (in nuclear physics,
this method is commonly named Green's function Monte Carlo). 
The DMC results of Ceperley and Alder from 1980 
\cite{ceperley1980} are still, more than thirty years later, 
among the most accurate energy estimates for the 
three-dimensional electron gas. In Ref.~\cite{ceperley1980},
the fermion sign problem is handled using 
a released-node approximation. Similar accuracy has been 
obtained in more recent calculations using the 
backflow-correlation technique \cite{kwon1998}, which is used 
to relax the simpler fixed-node approximation.

When modelling the electron gas using a finite box, as is
commonly done in Monte Carlo methods, the energy has an error
compared to the electron gas at the thermodynamic limit. 
At the thermodynamic limit, plane-wave single-particle states
fill the Fermi sphere with a continuous spectrum. When the
system is approximated using a finite box, the 
single-particle spectrum becomes discrete. This 
discretization gives an 
error that is common to systems with short-ranged and 
long-ranged interactions. The perhaps most obvious way to
correct for finite-size effects related to a discrete 
single-particle basis, is by using an extrapolation formula.
As is described in Ref.~\cite{drummond2008}, results from
Hartree-Fock and density-functional theory calculations 
can be used to construct extrapolation methods. Another
approach that efficiently reduces the finite-size error
is a technique that utilizes so-called twisted boundary 
conditions \cite{lin2001}. Twisted boundary conditions 
means that the single-particle wave function is 
multiplied by a complex phase factor when moving from one 
simulation cell to a neighbouring cell. When averaging
over results obtained with different twist angles, the
energy estimates become much more accurate than when
using periodic boundary conditions \cite{lin2001}.

Finite-size approximations of extended Coulombic systems, 
such as the electron gas, have additional errors that are 
caused by the long-ranged interaction \cite{fraser1996,chiesa2006,drummond2008}. 
Interactions with electrons in
neighbouring cells can be summed using, for example,
Ewald's method \cite{ewald1921,fraser1996,wood2004,drummond2008}. In 
Ewald's approach, each electron in the simulation cell
interacts with an infinite number of image charges located
at the same local position in all other cells. As is shown 
in Refs.~\cite{fraser1996,chiesa2006}, 
Ewald's interaction cannot describe the exchange-correlation
energy correctly. Chiesa \emph{et al.} \cite{chiesa2006}
used a static structure factor and a Jastrow factor
derived from RPA to estimate the correction to Ewald's
method. The correction technique of Chiesa \emph{et al.}
is directly applicable only to Monte Carlo methods.
Fraser \emph{et al.} \cite{fraser1996} suggested two 
different effective interactions that avoid the screening 
effects caused by Ewald's interaction . The most successful
alternative was to use a normal Coulomb interaction combined 
with the minimum-image convention \cite{fraser1996}.
The reader is referred to Ref.~\cite{drummond2008} for
more details about finite-size effects and different
approaches to correct for the related errors.

The full configuration-interaction quantum Monte Carlo 
(FCIQMC) met\-hod \cite{booth2009} is a new approach to the 
quantum many-body problem, in which the full 
configuration-interaction (FCI) equations \cite{harris1992} are 
solved approximately using a Monte Carlo technique. Similarly 
as in the DMC or GFMC methods, the Schr{\"o}dinger equation is 
written as a diffusion equation with an imaginary time 
variable \cite{shepherd2012a}. In the FCI method, the total 
wave-function ansatz 
is expressed as a linear combination of Slater determinants 
constructed from a given single-particle basis. In FCIQMC, 
the coefficients in the expansion of Slater determinants 
are obtained as the large-time limit of a random walk 
\cite{shepherd2012a}. As we discussed above, Monte Carlo
calculations of the electron gas have an error related to 
the finite number of particles in the simulation cell. 
Methods such as FCIQMC, CC, and partial summations derived 
from many-body perturbation theory have and additional error 
when studying systems with a discrete single-particle basis: 
The result depends on the given set of both occupied and 
unoccupied single-particle states \cite{shepherd2012b}. 
This error can be corrected for by using, for example, the 
single-point extrapolation technique \cite{shepherd2012b,
shepherd2012c} introduced by Shepherd \emph{et al.} The 
FCIQMC method has recently been applied to finite 
electron-gas systems \cite{shepherd2012a,shepherd2012b,
shepherd2012c}, giving results in close agreement
with DMC calculations utilizing backflow correlations 
\cite{shepherd2012a}.

Singal and Das \cite{singal1973} were the first to study 
the electron gas using a CC approach. The approximation
they used is similar to the BHF method \cite{haftel1970}, 
and does not properly include particle-hole ring diagrams.
Later, Freeman did CC calculations \cite{freeman1977} in 
which only ring diagrams and their exchange parts were 
retained in a CC doubles approximation. The results of both
Singal and Das \cite{singal1973} and Freeman 
\cite{freeman1977} compared well with dielectric-function
approaches. Presently, the most accurate CC caclulations of 
the three-dimensional electron gas are those of Bishop and
L{\"u}hrmann \cite{bishop1978,bishop1982}. Bishop and
L{\"u}hrmann derived a CC SUB2 approximation (also called
CC doubles (CCD) \cite{shavittbartlett2009}) for the electron gas, 
extended with some ladder contributions from higher-order
amplitudes. As the authors describe in 
Ref.~\cite{bishop1978}, the CCD approximation 
contains many more diagrammatic classes than
partial-summation techniques derived from perturbation 
theory. The CCD approximation takes account of 
particle-particle and hole-hole ladders, particle-hole
ring diagrams including exchange terms, and many other
diagrams to infinite order in perturbation theory. 
In their CC approximations, Bishop and L{\"u}hrmann 
replaced summations over hole states by averages. In
fact, the nine-dimensional CC amplitudes were simplified
to one-dimensional objects, with the absolute value
of the transfer momentum being the only variable. 
The authors showed that the state-average approximation
is accurate in the RPA approximation.
Bishop and L{\"u}hrmann neglected some diagrammatic 
classes, such as the hole-hole ladders and mixed ladders, 
which they assumed to be small. Despite all these 
approximations, the final CC energies are very accurate 
in a typical metallic density range. Relative 
differences of less than one percent
compared to DMC calculations of Ceperley and Alder
may indicate that the extended CCD approximation describes
most of the relevant correlations in the three-dimensional
electron gas \cite{bishop1982}.

The CC study of Bishop and L{\"u}hrmann was very successful, but still
the calculations are based on a large number of approximations. It
would be desireable to apply CC theory to the electron gas without the
same approximations.  Recently, Shepherd \emph{et al.}
\cite{shepherd2012b,shepherd2013a,shepherd2013b,shepherd2013c} have
taken up again the CC effort for the three-dimensional electron
gas. The CC calculations for the electron gas from the 1970s and 1980s
were all done at the thermodynamic limit
\cite{singal1973,freeman1977,bishop1978,bishop1982}. Unfortunately,
singularities related to the Coulomb two-body interaction may
complicate numerical calculations at the thermodynamic limit.
Shepherd \emph{et al.} approximate the electron gas using finite cubic
boxes
\cite{shepherd2012b,shepherd2013a,shepherd2013b,shepherd2013c}. When
using finite-size systems, the CC equations are simpler and most of
the problems with ill-bahaved functions are avoided.  Instead, one has
to deal with errors related to finite particle numbers and
single-particle bases, similarly as in the FCIQMC method
\cite{shepherd2012a,shepherd2012b,shepherd2012c}.

Let us finally turn our attention to the two-dimensional
electron gas. The two-dimensional electron gas is defined
in the same way as the three-dimensional counterpart, and 
similar approaches can therefore often be used to 
study both systems. As a first example, the classical 
derivation of the high-density RPA approximation by Gell-Mann 
and Brueckner \cite{gellmann1957} has been extended to the 
two-dimensional electron gas by Rajagopal and Kimball 
\cite{rajagopal1977}.

As far as we know, the only CC calculations that have been
done for the two-dimensional electron gas are the ring
\cite{freeman1978} and particle-particle ladder \cite{freeman1983}
approximations of Freeman. Beacuse of stronger correlations
in the purely two-dimensional system, the ring approximation
was not as reliable in two dimensions as it was in the 
three-dimensional case \cite{freeman1978}. According to the 
studies of Freeman, the CC ring-diagram approximation is 
less accurate than the CC ladder approximation at 
intermediate and low densities, whereas both methods give 
similar results at the high-density limit \cite{freeman1983}.

In the earliest CC studies of the 
electron gas \cite{singal1973,freeman1977,bishop1978,
freeman1978,bishop1982,freeman1983}, the system was treated 
at the thermodynamic limit. 
Some of the most accurate calculations of the electron gas
\cite{ceperley1980,tanatar1989,shepherd2012a} have been done using finite boxes and Monte
Carlo methods. Recently, CC theory has been applied in 
several studies \cite{shepherd2012b,shepherd2013a,
shepherd2013b,shepherd2013c,roggero2013} to finite-particle 
approximations of the homogeneous electron gas. 
 Freeman has done CC studies of the two-dimensional 
electron gas including only ring \cite{freeman1978} or 
only particle-particle ladder \cite{freeman1983} diagrams 
from the CCD approximation, but to the best of our knowledge, 
nobody has done a complete CCD calculation of the 
two-dimensional homogeneous electron gas. 
As an extension of the CC calculations by Shepherd 
\emph{et al.}~\cite{shepherd2012b,shepherd2013a,shepherd2013b,shepherd2013c}
and Roggero \emph{et al.}~\cite{roggero2013}, we study
the two-dimensional electron gas in the CCD approximation
using a finite number of particles.
We validate our methods by comparing results 
for the three-dimensional electron gas with Shepherd 
\emph{et al.} \cite{shepherd2012b,shepherd2012c}.
using momentum discretized in cartesian coordinates, see discussions below.

\subsection{Model Hamiltonian and Definitions}

The homogeneous electron gas is defined as an infinite system
consisting of electrons distributed with a constant density 
through the entire real space. The system is assumed to 
contain a constant positive background charge which cancels
the negative charges of the electrons. The Hamiltonian 
operator of the homogeneous electron gas can be written as
\cite{fetter}
\begin{align}
  \hat{H} = \hat{H}_{\text{kin}} + \hat{H}_{ee} + \hat{H}_{eb} 
  + \hat{H}_{bb},
\end{align}
where $\hat{H}_{\text{kin}}$ is the kinetic-energy operator,
$\hat{H}_{ee}$ models the electron-electron interaction,
$\hat{H}_{eb}$ represents the interaction between electrons 
and the positive background charge, and the operator 
$\hat{H}_{bb}$ gives the interaction energy of the background
charge with itself.

Similarly as in
Refs.~\cite{shepherd2012b,shepherd2013a,shepherd2013b,shepherd2013c,roggero2013},
we approximate the homogeneous electron gas by a finite hypercube
filled with electrons. The summation over
an infinite number of Coulomb interaction terms can be efficiently
calculated using, for example, Ewald's summation technique
\cite{ewald1921,fraser1996}.  In Ewald's approach, the Coulomb
interaction is split into a short-ranged part, which is calculated
in real space, and a long-ranged part, which is summed in Fourier
space.  The kinetic energy operator is
\begin{align}
  \hat{H}_{\text{kin}} = -\frac{\hbar^{2}}{2m}\sum_{i=1}^{N}
  \nabla_{i}^{2},
\end{align}
where the sum is taken over all particles in the finite
box. The Ewald electron-electron interaction operator 
can be written as \cite{drummond2008}
\begin{align}
  \hat{H}_{ee} = \sum_{i<j}^{N} 
  v_{E}\left( \mathbf{r}_{i}-\mathbf{r}_{j}\right)
  + \frac{1}{2}Nv_{0},
\end{align}
where $v_{E}(\mathbf{r})$ is the effective two-body 
interaction and $v_{0}$ is the self-interaction, defined 
as $v_{0} = \lim_{\mathbf{r} \rightarrow 0} 
\left\{ v_{E}(\mathbf{r}) - 1/r\right\} $. The negative 
electron charges are neutralized by a positive, homogeneous 
background charge. Fraser \emph{et al.} explain 
\cite{fraser1996} how the
electron-background and background-background terms, 
$\hat{H}_{eb}$ and $\hat{H}_{bb}$, vanish
when using Ewald's interaction for the three-dimensional
electron gas. Using the same arguments, one can show that
these terms are also zero in the corresponding 
two-dimensional system.

In the three-dimensional electron gas, the Ewald 
interaction is \cite{drummond2008}
\begin{align}
  v_{E}(\mathbf{r}) &= \sum_{\mathbf{k} \neq \mathbf{0}}
  \frac{4\pi }{L^{3}k^{2}}e^{i\mathbf{k}\cdot \mathbf{r}}
  e^{-\eta^{2}k^{2}/4} \nonumber \\
  &+ \sum_{\mathbf{R}}\frac{1}{\left| \mathbf{r}
    -\mathbf{R}\right| } \erfc \left( \frac{\left| 
    \mathbf{r}-\mathbf{R}\right|}{\eta }\right)
  - \frac{\pi \eta^{2}}{L^{3}},
\end{align}
where $L$ is the box side length, $\erfc(x)$ is the 
complementary error function, and $\eta $ is a free
parameter that can take any value in the interval 
$(0, \infty )$. The translational vector 
\begin{align}
  \mathbf{R} = L\left(n_{x}\mathbf{u}_{x} + n_{y}
  \mathbf{u}_{y} + n_{z}\mathbf{u}_{z}\right) ,
\end{align}
where $\mathbf{u}_{i}$ is the unit vector for dimension $i$,
is defined for all integers $n_{x}$, $n_{y}$, and 
$n_{z}$. These vectors are used to obtain all image
cells in the entire real space. 
The parameter $\eta $ decides how 
the Coulomb interaction is divided into a short-ranged
and long-ranged part, and does not alter the total
function. However, the number of operations needed
to calculate the Ewald interaction with a desired 
accuracy depends on $\eta $, and $\eta $ is therefore 
often chosen to optimize the convergency as a function
of the simulation-cell size \cite{drummond2008}. In
our calculations, we choose $\eta $ to be an infinitesimally
small positive number, similarly as was done in 
Refs.~\cite{shepherd2012b,shepherd2013a,shepherd2013b,shepherd2013c,roggero2013}.
This gives an interaction that is evaluated only in
Fourier space.

When studying the two-dimensional electron gas, we
use an Ewald interaction that is quasi two-dimensional.
The interaction is derived in three dimensions, with 
Fourier discretization in only two dimensions 
\cite{wood2004}. The Ewald effective
interaction has the form \cite{wood2004}
\begin{align}
  v_{E}(\mathbf{r}) &= \sum_{\mathbf{k} \neq \mathbf{0}} 
  \frac{\pi }{L^{2}k}\left\{ e^{-kz} \erfc \left(
  \frac{\eta k}{2} - \frac{z}{\eta }\right)+ \right. \nonumber \\
  & \left. e^{kz}\erfc \left( \frac{\eta k}{2} + \frac{z}{\eta }
  \right) \right\} e^{i\mathbf{k}\cdot \mathbf{r}_{xy}} 
  \nonumber \\
  & + \sum_{\mathbf{R}}\frac{1}{\left| \mathbf{r}-\mathbf{R}
    \right| } \erfc \left( \frac{\left| \mathbf{r}-\mathbf{R}
    \right|}{\eta }\right) \nonumber \\
  & - \frac{2\pi}{L^{2}}\left\{ z\erf \left( \frac{z}{\eta }
  \right) + \frac{\eta }{\sqrt{\pi }}e^{-z^{2}/\eta^{2}}\right\},
\end{align}
where the Fourier vectors $\mathbf{k}$ and the position vector
$\mathbf{r}_{xy}$ are defined in the $(x,y)$ plane. When
applying the interaction $v_{E}(\mathbf{r})$ to two-dimensional
systems, we set $z$ to zero. Similarly as in the 
three-dimensional case, and as suggested in 
Ref.~\cite{drummond2008} for two dimensions, also here we 
choose $\eta $ to approach zero from above. The resulting 
Fourier-transformed interaction is
\begin{align}
  v_{E}^{\eta = 0, z = 0}(\mathbf{r}) = \sum_{\mathbf{k} \neq \mathbf{0}} 
  \frac{2\pi }{L^{2}k}e^{i\mathbf{k}\cdot \mathbf{r}_{xy}}. 
\end{align}
The self-interaction $v_{0}$ is a constant that can be 
included in the reference energy.

In the three-dimensional electron gas, the antisymmetrized
matrix elements are
\begin{align} \label{eq:vmat_3dheg}
  & \langle \mathbf{k}_{p}m_{s_{p}}\mathbf{k}_{q}m_{s_{q}}
  |\tilde{v}|\mathbf{k}_{r}m_{s_{r}}\mathbf{k}_{s}m_{s_{s}}\rangle_{AS} 
  \nonumber \\
  & = \frac{4\pi }{L^{3}}\delta_{\mathbf{k}_{p}+\mathbf{k}_{q},
    \mathbf{k}_{r}+\mathbf{k}_{s}}\left\{ 
  \delta_{m_{s_{p}}m_{s_{r}}}\delta_{m_{s_{q}}m_{s_{s}}}
  \left( 1 - \delta_{\mathbf{k}_{p}\mathbf{k}_{r}}\right) 
  \frac{1}{|\mathbf{k}_{r}-\mathbf{k}_{p}|^{2}}
  \right. \nonumber \\
  & \left. - \delta_{m_{s_{p}}m_{s_{s}}}\delta_{m_{s_{q}}m_{s_{r}}}
  \left( 1 - \delta_{\mathbf{k}_{p}\mathbf{k}_{s}} \right)
  \frac{1}{|\mathbf{k}_{s}-\mathbf{k}_{p}|^{2}} 
  \right\} ,
\end{align}
where the Kronecker delta functions 
$\delta_{\mathbf{k}_{p}\mathbf{k}_{r}}$ and
$\delta_{\mathbf{k}_{p}\mathbf{k}_{s}}$ ensure that the 
contribution with zero momentum transfer vanishes.
Similarly, the matrix elements for the two-dimensional
electron gas are
\begin{align} \label{eq:vmat_2dheg}
  & \langle \mathbf{k}_{p}m_{s_{p}}\mathbf{k}_{q}m_{s_{q}}
  |v|\mathbf{k}_{r}m_{s_{r}}\mathbf{k}_{s}m_{s_{s}}\rangle_{AS} 
  \nonumber \\
  & = \frac{2\pi }{L^{2}}
  \delta_{\mathbf{k}_{p}+\mathbf{k}_{q},\mathbf{k}_{r}+\mathbf{k}_{s}}
  \left\{ \delta_{m_{s_{p}}m_{s_{r}}}\delta_{m_{s_{q}}m_{s_{s}}} 
  \left( 1 - \delta_{\mathbf{k}_{p}\mathbf{k}_{r}}\right)
  \frac{1}{
    |\mathbf{k}_{r}-\mathbf{k}_{p}|} \right.
  \nonumber \\
  & - \left. \delta_{m_{s_{p}}m_{s_{s}}}\delta_{m_{s_{q}}m_{s_{r}}}
  \left( 1 - \delta_{\mathbf{k}_{p}\mathbf{k}_{s}}\right)
  \frac{1}{ 
    |\mathbf{k}_{s}-\mathbf{k}_{p}|}
  \right\} ,
\end{align}
where the single-particle momentum vectors $\mathbf{k}_{p,q,r,s}$
are now defined in two dimensions. In our calculations, the 
self-interaction constant is included in the reference 
energy, as in Eq.~(\ref{eq:ene_ref}). We therefore get the 
Fock-operator matrix elements 
\begin{align}
  \langle \mathbf{k}_{p}|f|\mathbf{k}_{q} \rangle
  = \frac{\hbar^{2}k_{p}^{2}}{2m}\delta_{\mathbf{k}_{p},
  \mathbf{k}_{q}} + \sum_{\mathbf{k}_{i}}\langle 
  \mathbf{k}_{p}\mathbf{k}_{i}|v|\mathbf{k}_{q}
  \mathbf{k}_{i}\rangle_{AS}.
  \label{eq:fock_heg}
\end{align}
In Ref.~\cite{shepherd2013b}, the matrix elements were 
defined with the self-interaction constant included in the
two-body interaction. This gives Fock-operator matrix 
elements with a gap constant. The definition used in 
Ref.~\cite{shepherd2013b} may give numerically more
stable calculations, as the gap constant 
prevents the energy denominator from becoming too small
in the vicinity of the Fermi surface. However, when using
Fock matrix elements as defined in Eq.~(\ref{eq:fock_heg}), 
the energy denominator in the CC equations never vanishes 
unless the numerator is zero. 
When using periodic boundary conditions, the 
discrete-momentum single-particle basis functions 
\[
\phi_{\mathbf{k}}(\mathbf{r}) =
e^{i\mathbf{k}\cdot \mathbf{r}}/L^{d/2}
\]
are associated with 
the single-particle energy   
\begin{align}
  \varepsilon_{n_{x}, n_{y}} = \frac{\hbar^{2}}{2m} 
  \left( \frac{2\pi }{L}\right)^{2}
  \left( n_{x}^{2} + n_{y}^{2}\right)
\end{align}
for two-dimensional sytems and 
\begin{align}
  \varepsilon_{n_{x}, n_{y}, n_{z}} = \frac{\hbar^{2}}{2m}
  \left( \frac{2\pi }{L}\right)^{2}
  \left( n_{x}^{2} + n_{y}^{2} + n_{z}^{2}\right)
\end{align} 
for three-dimensional systems. Similarly as in, for example, 
Refs.~\cite{shepherd2012a,roggero2013,hagen2014}, we choose 
the single-particle basis such that both the occupied and 
unoccupied single-particle spaces have a closed-shell 
structure. This means that all single-particle states 
corresponding to energies below a chosen cutoff are
included in the basis. We study only the unpolarized spin
phase, in which all orbitals are occupied with one spin-up 
and one spin-down electron. Table \ref{tab:orbitals_2dheg}
shows the lowest-lying spin-orbitals and the cumulated
numbers of single-particle states for a two-dimensional
electron box with periodic boundary conditions.
\begin{table}
  \begin{center}
  \begin{tabular}{c|c|c|c|c}
    $n_{x}^{2}+n_{y}^{2}$ & $n_{x}$ & $n_{y}$ & 
    $N_{\uparrow \downarrow }$ & $N_{\uparrow \uparrow }$ \\
    \hline
    0 & 0 & 0 & 2 & 1 \\
    \hline
    1 & -1 &  0 & & \\
    &  1 &  0 & & \\
    &  0 & -1 & & \\
    &  0 &  1 & 10 & 5 \\
    \hline
    2  & -1 & -1 & & \\
    & -1 &  1 & & \\
    &  1 & -1 & & \\
    &  1 &  1 & 18 & 9 \\
    \hline
    4  & -2 & 0 & & \\
    &  2 & 0 & & \\
    &  0 & -2 & & \\
    &  0 &  2 & 26 & 13 \\
    \hline
    5  & -2 & -1 & & \\
    &  2 & -1 & & \\
    & -2 &  1 & & \\
    &  2 &  1 & & \\
    & -1 & -2 & & \\
    & -1 &  2 & & \\
    &  1 & -2 & & \\
    &  1 &  2 & 42 & 21 \\
    \hline
  \end{tabular}
  \end{center}
  \caption{Illustration of how single-particle energies
    fill energy shells in a two-dimensional electron box.
  Here $n_{x}$ and $n_{y}$ are the momentum quantum numbers,
  $n_{x}^{2} + n_{y}^{2}$ determines the single-particle 
  energy level, $N_{\uparrow \downarrow }$ represents the 
  cumulated number of spin-orbitals in an unpolarized spin
  phase, and $N_{\uparrow \uparrow }$ stands for the
  cumulated number of spin-orbitals in a spin-polarized
  system.} 
  \label{tab:orbitals_2dheg}
\end{table}

Finally, a useful benchmark for our calculations is the expression for
the reference energy $E_0$ per particle.
Defining the $T=0$ density $\rho_0$, we can in turn determine in three
dimensions the radius $r_0$ of a sphere representing the volume an
electron occupies (the classical electron radius) as
\[
r_0= \left(\frac{3}{4\pi \rho}\right)^{1/3}.
\]
In two dimensions the corresponding quantity is
\[
r_0= \left(\frac{1}{\pi \rho}\right)^{1/2}.
\]
One can then express the reference energy per electron in terms of the
dimensionless quantity $r_s=r_0/a_0$, where we have introduced the
Bohr radius $a_0=\hbar^2/e^2m$. The energy per electron computed with
the reference Slater determinant can then be written as \cite{fetter}
(using hereafter only atomic units, meaning that $\hbar = m = e = 1$)
\[
E_0/N=\frac{1}{2}\left[\frac{2.21}{r_s^2}-\frac{0.916}{r_s}\right],
\]
for the three-dimensional electron gas.  For the two-dimensional gas
the corresponding expression is, see for example
Refs.~\cite{rajagopal1977,tanatar1989},
\[
E_0/N=\frac{1}{r_s^2}-\frac{8\sqrt{2}}{3\pi r_s}.
\]

For an infinite homogeneous system, there are some particular
simplications due to the conservation of the total momentum of the
particles.  By symmetry considerations, the total momentum of the
system has to be zero zero. Both the kinetic energy operator and the
total Hamiltonian $\hat{H}$ are assumed to be diagonal in the total
momentum $\mathbf{K}$. Hence, both the reference state $\Phi_{0}$ and
the correlated ground state $\Psi$ must be eigenfunctions of the
operator $\mathbf{\hat{K}}$ with the corresponding eigenvalue
$\mathbf{K} = \mathbf{0}$ \cite{day1967}.  This leads to important
simplications to our different many-body methods. In coupled cluster
theory for example, see the discussion in the next subsection, all
terms that involve single particle-hole excitations vanish.

\subsection{Coupled-cluster theory for the homogeneous electron gas}

In this subsection we present the coupled-cluster equations needed to
obtain the correlation energy at the singles and doubles level. Due to
conservation of momentum, the coupled cluster equations simplify
considerably.  The so-called singles equa

The foundation for most many-body methods is to express the correct
wave function by an expansion in a set of basis functions.  One
example is the Hartree-Fock~(HF) method which employs a unitary
transformation of the single-particle wave functions,
\begin{equation}
|\lambda \rangle = \sum_{\psi} C_{\lambda \psi} |\psi \rangle ,
\end{equation}
and approximates the ground state with a reference Slater determinant
built up by these transformed wave functions.  Another example is
configuration interaction~(CI) where the reference determinant is set
to a linear expansion of determinants, including the initial reference
determinant, 1p-1h excitations, 2p-2h excitations and so on, i.e.
\begin{equation}
|\Psi_{0}^{CI} \rangle = C_0 |\Phi_0 \rangle + \sum_{ia} C_i^a
|\Phi_i^a\rangle + \sum_{ijab} C_{ij}^{ab} |\Phi_{ij}^{ab} \rangle +
\cdots \hspace{2mm}.
\end{equation}
In all these methods one needs to solve a set of coupled equations to
find the coefficients.

The coupled-cluster method also expands the exact solution in a set of
Slater determinants, but employs a non-linear expansion through the
exponential ansatz,
\begin{equation}
\label{eq:CC:expon}
|\Psi_0^{CC} \rangle = e^{\hat{T}} |\Phi_0 \rangle ,
\end{equation}
where $\hat{T}$ is the cluster operator including \textit{all}
possible excitations on the reference determinant.  Sorting
excitations by the number of excited electrons, we may generally
express this general cluster operator as a sum of a 1p-1h operator, a
2p-2h operator, and so on,
\begin{equation}
\hat{T} = \hat{T_1} + \hat{T_2} + \hat{T_3} + \cdots \hspace{2mm}.
\end{equation}
In the form of second-quantized operators the 1p-1h cluster operator
is defined as
\begin{equation}
\hat{T}_1 = \sum_{ia} t_i^a \hat{a}^{\dagger} \hat{i},
\end{equation}
the 2p-2h cluster operator as
\begin{equation}
\hat{T}_2 = \frac{1}{4} \sum_{ijab} t_{ij}^{ab} \hat{a}^{\dagger}
\hat{b}^{\dagger} \hat{j} \hat{i},
\end{equation}
continuing up to
\begin{equation}
\hat{T}_n = \left( \frac{1}{n!}\right)^2 \sum_{ij\cdots ab\cdots}
t_{ij\cdots}^{ab\cdots} \hat{a}^{\dagger} \hat{b}^{\dagger} \cdots
\hat{j} \hat{i} .
\end{equation}
As long as we have a complete single-particle basis and include all
possible excitations up to $n$p-$n$h in a system with $n$ particles,
we should find the exact solution for both CI, $|\Psi_0^{CI}\rangle$,
and CC, $|\Psi_0^{CC}\rangle$.

Adding all terms together we get the complete $\hat{T}_1$ amplitude
equations;
\begin{equation}
\label{eq:CC:t1eq_raw}
\begin{split}
0 =& f_{ai} + \sum_d f_{ad} t_i^d - \sum_l f_{li} t_l^a + \sum_{ld}
\langle la||di \rangle t_l^d \\ &+ \sum_{ld} f_{ld} t_{il}^{ad} +
\frac{1}{2} \sum_{lde} \langle al||de \rangle t_{il}^{de} -
\frac{1}{2} \sum_{lmd} \langle lm||di \rangle t_{lm}^{da} - \sum_{ld}
f_{ld} t_i^d t_l^a \\ & + \sum_{lde} \langle al||de \rangle t_i^d
t_l^e - \sum_{lmd} \langle lm||di \rangle t_l^d t_m^a + \frac{1}{2}
\sum_{lmde} \langle lm||de \rangle t_i^d t_{lm}^{ea} \\ &+ \frac{1}{2}
\sum_{lmde} \langle lm||de \rangle t_m^a t_{il}^{de} + \sum_{lmde}
\langle lm||de \rangle t_m^e t_{il}^{ad} + \sum_{lmde} \langle lm||de
\rangle t_i^d t_l^e t_m^a .
\end{split}
\end{equation}
\begin{equation}
\label{eq:CC:t2eq_raw}
\begin{split}
0 =& \langle ab || ij \rangle +\frac{1}{2} \langle ab||de \rangle
t_{ij}^{de} \\ & - \hat{P}_{ij} \left[ f_{li} + f_{ld} t_{i}^d +
  \langle ml||di \rangle t_m^d + \langle ml||de \rangle t_m^d t_i^e
  +\frac{1}{2}\langle ml||de \rangle t_{mi}^{de} \right] t_{lj}^{ab}
\\ & + \frac{1}{2} \left[ \langle lm||ij \rangle + \hat{P}_{ij}
  \langle lm||dj \rangle t_i^d +\frac{1}{2} \langle lm||de \rangle
  t_{ij}^{de} +\hat{P}_{ij} \frac{1}{2} \langle lm||de \rangle t_i^d
  t_j^e \right] t_{lm}^{ab} \\ & + \hat{P}_{ab} \left[ f_{bd} - f_{ld}
  t_l^b + \langle bl||de \rangle t_l^e - \langle lm||de \rangle t_m^e
  t_l^b + \frac{1}{2} \langle lm||de \rangle t_{lm}^{eb} \right]
t_{ij}^{ad} \\ & +\hat{P}_{ij} \hat{P}_{ab} \left[ \langle lb||dj
  \rangle - \langle lm||dj \rangle t_{m}^b + \langle bl||ed \rangle
  t_{j}^e - \langle lm||de \rangle t_j^e t_m^b + \frac{1}{2} \langle
  lm||de \rangle t_{mj}^{eb} \right] t_{il}^{ad} \\ & -\hat{P}_{ab}
\left[ \langle al||ij \rangle + \frac{1}{2} \langle al||de \rangle
  t_{ij}^{de} + \hat{P}_{ij} \langle al||dj \rangle t_i^d +
  \hat{P}_{ij} \frac{1}{2} \langle al||de \rangle t_i^d t_j^e \right.
  \\& \left. + \frac{1}{2} \langle lm||ij \rangle t_m^a + \frac{1}{4}
  \langle lm||de \rangle t_m^a t_{ij}^{de} + \hat{P}_{ij} \frac{1}{2}
  \langle lm||dj \rangle t_i^d t_m^a + \hat{P}_{ij} \frac{1}{4}
  \langle lm||de \rangle t_i^d t_j^e t_m^a \right] t_l^b \\ & +
\hat{P}_{ij} \left[ \langle ab || dj \rangle + \frac{1}{2} \langle
  ab||de \rangle t_j^e \right] t_i^d .
\end{split}
\end{equation}

\subsection{The Similarity Renormalization Group flow equations}
\label{subsec:SRG}
The Similarity Renormalization Group~(SRG) method was introduced
independently by Glazek and Wilson
\cite{PhysRevD.48.5863,*PhysRevD.49.4214} and Wegner
\cite{PhysRepWegner0,*PhysRepWegner} as a new way to implement the
principle of energy scale separation.  The method uses a continuous
sequence of unitary transformations to decouple the high- and
low-energy matrix elements of a given interaction, thus driving the
Hamiltonian towards a band- or block-diagonal form.

\section{Results}
\label{sec:results}
\subsection{Benchmark results for the three-dimensional electron gas}

\subsection{The two-dimensional electron gas}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneN10Rs10.pdf}
    \end{center}
    \caption{Correlation energy per particle as function of the number of
single-particle orbits for a ten-particle electron gas at $r_s=1.0$, using CCD, IMSRG(2) 
and FCIQMC. The IMSRG(2)
correlation energies are in excellent agreement with the almost
exact FCIQMC results. For comparison, we have marked the correlation
energy in the thermodynamic limit, as obtained by extrapolating GFMC
energies. The width of the horizontal line shows the statistical error of
the GFMC calculations.}
            \label{fig:N10}
\end{figure}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneN26Rs10.pdf}
    \end{center}
    \caption{Correlation energy per particle as function of the number of
single-particle orbits for a 26 particle electron gas at $r_s=1.0$, using CCD, IMSRG(2) 
and FCIQMC. Here we see a larger discrepancy between the IMSRG(2) results and the FCIQMC values. 
For comparison, we have marked the correlation
energy in the thermodynamic limit, as obtained by extrapolating GFMC
energies.}
            \label{fig:N26}
\end{figure}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneNRs01.pdf}
    \end{center}
    \caption{CCD correlation energy per particle as function of the number of
single-particle orbits for varying  electron numebrs $N$ $r_s=0.1$.}
            \label{fig:ccd01}
\end{figure}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneNRs05.pdf}
    \end{center}
    \caption{CCD correlation energy per particle as function of the number of
single-particle orbits for varying  electron numebrs $N$ $r_s=0.5$.}
            \label{fig:ccd05}
\end{figure}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneNRs10.pdf}
    \end{center}
    \caption{CCD correlation energy per particle as function of the number of
single-particle orbits for varying  electron numebrs $N$ $r_s=1.0$.}
            \label{fig:ccd05}
\end{figure}

\begin{figure}[hbtp]
     \begin{center}

            \includegraphics[width=0.4\textwidth]{figures/nOrbitsEneNRs20.pdf}
    \end{center}
    \caption{CCD correlation energy per particle as function of the number of
single-particle orbits for varying  electron numebrs $N$ $r_s=2.0$.}
            \label{fig:ccd05}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
%
\begin{acknowledgments}
  This work was supported by the

\end{acknowledgments}

\bibliography{paper} \bibliographystyle{apsrev4-1}

\end{document}

  \begin{itemize}
  \item The first study of the 2D electron gas with SRG, FCIQMC, and CC  
  \item How well can the methods describe correlations?
  \item At what densities do SRG and CC break down?
  \item Theoretical explanation of breakdown?
  \item Ring and ladder approximations vs. CCD 
  \item What are differences between finite and periodic systems
    (QD vs. 2DEG)
  \item Compare 2D with finite thickness 2D+1D (or 3D?)
  \item Virial theorem (J{\o}rgen's thesis)?
  \item More ideas?
  \end{itemize}