### 2019-04-30 doc/PhysicsToday/PhysToday.tex
added physics today material

\documentclass[reprint,floatfix]{revtex4-1}

\usepackage{graphicx}
\usepackage{comment}
\usepackage{lipsum}
\usepackage[utf8]{inputenc}
\usepackage[toc,page]{appendix}
\usepackage{braket}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{pgfplots}
\usepackage{bbold}

\begin{document}

\title{Integrating a Computational Perspective in Physics Education}

\author{D.M. Caballero}
\email{haakoek@uio.no}
\affiliation{National Superconducting Cyclotron Laboratory and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Morten Hjorth-Jensen}
\email{hjensen@msu.edu}
\affiliation{National Superconducting Cyclotron Laboratory and Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}
\date{\today}

\begin{abstract}
    This is the abstract.
\end{abstract}

\maketitle

\bibliographystyle{ieeetr}
\bibliography{refs}
\end{document}